---
title: "One Sample T-Test Against Zero"
author: "Bryant Byrd"
format: html
editor: visual
---

## Study 4 Thesis 

\# Purpose\
This analysis tests whether the mean percent difference between IPC and Sham performance is significantly different from zero.\
A percent difference of zero would mean there was no average difference between IPC and Sham.\
A negative value means IPC was faster than Sham.\
A positive value means IPC was slower than Sham.

```{r}
# Load packages

library(tidyverse)

library(readxl)

performance_outcome <- read_excel("data/performance_outcome.xlsx")

# Clean the data

matched_long <- performance_outcome %>%
  rename(
    pair_id = ID,
    race = Race,
    trial = Trial,
    finish_time_excel = `Finsished Time`,
    finish_sec = Finished_Sec,
    pct_diff_excel = `%diff`,
    place = Place
  ) %>%
  filter(!is.na(pair_id), trial %in% c("IPC", "Sham"))
```

```{r}
# Convert to one row per matched pair

matched_wide <- matched_long %>%
  select(pair_id, race, trial, finish_sec, place) %>%
  pivot_wider(
    names_from = trial,
    values_from = c(race, finish_sec, place),
    names_sep = "_"
  ) %>%
  mutate(
    pct_diff_decimal = (finish_sec_IPC - finish_sec_Sham) / finish_sec_Sham,
    pct_diff_percent = pct_diff_decimal * 100
  )

matched_wide
```

```{r}
# One-sample t-test against zero

t_test_result <- t.test(
  matched_wide$pct_diff_percent,
  mu = 0
)

t_test_result
```

\# Interpretation\
The one-sample t-test was used to determine whether the average percent difference between IPC and Sham was significantly different from zero.\
\
If the p-value is less than 0.05, the mean percent difference is statistically significantly different from zero.\
\
If the mean percent difference is negative, IPC was faster than Sham on average.\
\
If the mean percent difference is positive, IPC was slower than Sham on average.
